---
title: "Assignment 4 - Canonical Correlation Analysis"
author: |
  | **GROUP 03**
  | Agustin Valencia - **aguva779**
  | Bayu Brahmantio  - **baybr878**
  | Joris van Doorn - **jorva845**
  | Marcos F Mourao - **marfr825**
date: "`r format(Sys.time(), '%d %B %Y')`"
output:
  pdf_document: default
---

```{r setup, include=FALSE, echo=FALSE}
RNGversion('3.5.1')
knitr::opts_chunk$set(echo = TRUE)
library(expm)
library(knitr)
library(CCP)
```

```{r, echo=FALSE}
data <- read.table("./Data/P10-16.DAT")
```

# Canonical correlation analysis by utilizing suit able software

Look at the data described in Exercise 10.16 of Johnson, Wichern. You may find it in the file P10-16.DAT. The data for 46 patients are summarized in a covariance matrix, which will be analyzed in R. Read through the description of the different R packages and functions so you may chose the must suitable one for the analysis. Supplement with own code where necessary.

```{r, echo=FALSE}
#number of observations (patients)
n <- 46
#number of primary variables
p <- 3
#number of secondary variables
q <- 2
```

The given matrix is the following: 

```{r, echo=FALSE}
kable(data)
```

Thus, separating the variance-covariance matrix it is obtained that

```{r, results=FALSE, echo=FALSE}
#separating the variance-covariance matrix
sigma11 <- as.matrix(data[1:3,1:3])
sigma22 <- as.matrix(data[4:5, 4:5])
sigma21 <- as.matrix(data[4:5, 1:3])
sigma12 <- as.matrix(data[1:3, 4:5])

M <- sqrtm(solve(sigma11)) %*% sigma12 %*% solve(sigma22) %*% sigma21 %*% sqrtm(solve(sigma11))
D <- sqrtm(solve(sigma22)) %*% sigma21 %*% solve(sigma11) %*% sigma12 %*% sqrtm(solve(sigma22))

```

$$\Sigma_{11} =$$
```{r, echo=FALSE}
kable(sigma11)
```

$$\Sigma_{22} =$$
```{r, echo=FALSE}
kable(sigma22)
```

$$\Sigma_{21} =$$
```{r, echo=FALSE}
kable(sigma21)
```

$$\Sigma_{12} =$$
```{r, echo=FALSE}
kable(sigma12)
```

It can be computed that

$$M = S_{11}^{-1/2}S_{12}S_{22}^{-1}S_{21}S_{11}^{-1/2} = $$
```{r, echo=FALSE}
kable(M)
```

Its eigen-decomposition is given by

```{r, echo=FALSE}
mEigenDecom <- eigen(M)
mEigVect <- mEigenDecom$vectors[,1:2]
mEigVals <- mEigenDecom$values[1:2]
```

The first two eigenvalues $\overrightarrow{\alpha}$:
```{r, echo=FALSE}
kable(mEigVals)
```

The first two eigenvectors $\mathbf{\overrightarrow{e}}$:
```{r, echo=FALSE}
kable(mEigVect)
```

Also, 

$$D = S_{22}^{-1/2}S_{21}S_{11}^{-1}S_{12}S_{22}^{-1/2} = $$

```{r, echo=FALSE}
kable(D)
```

Its eigen-decomposition is given by

```{r, echo=FALSE}
dEigenDecom <- eigen(D)
dEigVect <- dEigenDecom$vectors
dEigVals <- dEigenDecom$values
```

Eigenvalues $\overrightarrow{\beta}$:
```{r, echo=FALSE}
kable(dEigVals)
```

Eigenvectors $\mathbf{\overrightarrow{f}}$:
```{r, echo=FALSE}
kable(dEigVect)
```

It is defined that

$$\mathbf{\hat{a}'}_k = \mathbf{\hat{e}'}_k S_{11}^{-1/2}$$
$$\mathbf{\hat{b}'}_k = \mathbf{\hat{f}'}_k S_{22}^{-1/2}$$
Thus, based in results above: 

```{r, echo=FALSE}
a <- t(mEigVect) %*% sqrtm(solve(sigma11))
a <- t(a)
b <- t(dEigVect) %*% sqrtm(solve(sigma22))
b <- t(b)
```

$$\mathbf{\hat{a}} = $$

```{r, echo=FALSE}
kable(a)
```

$$\mathbf{\hat{b}} = $$
```{r, echo=FALSE}
kable(b)
```


$M$ and $D$ have the same eigenvalues, so the canonical correlations can be calculated from one of 

$$[\hat{\rho_1^*}, \hat{\rho_2^*} ]=$$
```{r, echo=FALSE}
rho1 <- sqrt(mEigVals)
rho2 <- sqrt(dEigVals)
kable(t(rho1))
```



a) **Test at the 5% level if there is any association between the groups of variables.**

$$H_0: \Sigma_{12} = 0$$ 
$$H1: \Sigma{12} \neq 0$$
 We have $p \cdot q = 3 * 2 = 6$ degrees of freedom. For $n = 46$, we use Bartletts approximation and our test statistic is: 

$$ -\left( n - 1 - \frac{1}{2}(p+q+1)\right) \ln \prod_{i=1}^2{(1-\hat{\rho_i^{*2}})} $$

```{r, echo=FALSE}
test1 <- -(n-1-(0.5*(p+q+1))) * log(prod(1-rho1^2))
#Hypothesis testing
alpha <- 0.05
#critical value
crit <- qchisq(p = (1-alpha), df = p*q)
#test statistic
test1 <- -(n-1-(0.5*(p+q+1))) * log(prod(1-rho1^2))  #13.74948
test2 <- -(n-1-(0.5*(p+q+1))) * log(1-rho1[2]^2)  #13.74948 they are the same
```

For the obtained canonical correlations it is obtained that

$$test_1 = 13.74948$$
$$test_2 = 0.6668632$$

It is seen that $test_1 > 12.59159$, thus according to $test_1$ $H_0$ should be rejected. Now, $test_2 < 12.59159$, then according to $test_2$ we fail to reject $H_0$.

In summary we conclude that we cannot assume that $$H_0: \Sigma_{12} = 0$$ at a significance level $\alpha=0.05$





b) **How many pairs of canonical variates are significant?**

$$H_0: \rho_2 \neq 0$$ and  $$H_1: \rho_2 = 0$$
At $\alpha = 0.05$ significance, we get a test statistic of 0.66, smaller than the critical value of 12.59. We then reject $H_0$, meaning only the first canonical correlation ($\rho_1$) is significantly different from 0.



c) **Interpret the "significant" squared canonical correlations.
Tip: Read section "Canonical Correlations as Generalizations of Other Correlation Coefficients".**

The significant squared canonical correlation ($\rho_1^2$) measures the overlap between $X^{(1)}$ and $X^{(2)}$, the different group of variables.


d) **Interpret the canonical variates by using the coefficients and suitable correlations.**

The canonical correlations variables are:
$$U_1  = 0.0131007 X_1^{(1)} - 0.0144383 X_2^{(1)} + 0.0233997 X_3^{(1)}$$
$$V_1 = -8.0655751 X_1^{(2)} + 0.0191591 X_2^{(2)}$$

We can interpret $U_1$ as a measure of how well a patient can process sugar (glucose) in their bloodstream. $V_1$, on the other hand, can be interpreted as how prone the patient is to diabetes.


e) **Are the "significant" canonical variates good summary measures of the respective data sets?
Tip: Read section "Proportions of Explained Sample Variance".**

```{r, echo=FALSE}
num <- t(a[,1]) %*% sigma12 %*% b[1,]
den <- sqrt(t(a[,1]) %*% sigma11 %*% a[,1]) * sqrt(t(b[,1]) %*% sigma22 %*% b[,1])
r <- num/den
print("The explained variance")
r
```

The correlation variate explains 44% of the total sample variance. We conclude therefore that the significant canonical variates are not a good summary because there might be nonlinear relations.


f) **Give your opinion on the success of this canonical correlation analysis.**

While the canonical correlation variables have insightful interpretration, the analysis is not good enough given the small summary capability of its variates.






\newpage
# Appendix A - Code

```{r, ref.label=knitr::all_labels(),echo=TRUE,eval=FALSE}
```
